---
title: "UCSF Stages Data Cleaning - Prep for Analysis"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

## Project 
MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description
This file imports demographic and staging data and returns a dataset indicating each patient's `max_stage`, `date_adm`, `date_disc`, `los`, `stage_adm`, `stage_disc`, `days_to_disc`. This file prepares the staging data for analysis. 

## Source Data
  - Demographics and Daily Covid Stage Data (`dm_covid_stg_11.08.2021.csv`)
    + This file contains data on patient demographics and COVID stage (based on WHO Clinical Progression Scale, which aims to capture patient clinical trajectory and resource usage over the course of the clinical illness -- in this case, COVID-19).
    + Each row represents one day per patient in hospital. 

# Load required packages
```{r Load required packages}
library(here)
library(tidyverse)
```

# Import and preview data 
```{r Import and preview data, echo=F}
df <- read_csv(here("data", "dm_covid_stg_11.08.21.csv"))

glimpse(df)

head(df, n=20)
```

# Create new variables: `max_stage`, `date_adm`, `date_disc`, `stage_adm`, `stage_disc`, `days_to_disc`, `DOD`, 
```{r Create new variables}
df <- df %>% 
  rename(stage = STAGE) %>%
  group_by(ID) %>% 
  mutate(max_stage = max(stage),                # max_stage
         date_adm = min(date),                  # date_adm
         date_disc = max(date),                 # date_disc
         stage_adm = stage[date == date_adm],   # stage_adm
         stage_disc = stage[date == date_disc], # stage_disc
         days_to_disc = as.integer(difftime(date_disc, date, units = 'days')),
                                                # days_to _disc
         DOD = date[DEATH == 1],                # date of death (DOD)
         sex = as.factor(sex),                  # sex as factor
         race = as.factor(race),                # race as factor
         ethnicity = as.factor(ethnicity),      # ethnicity as factor
         smoking = as.factor(smoking),          # smoking as factor
         death = as.factor(end_in_death)) %>%   # death as factor
  select(-c(DEATH, end_in_death))

# Check variables types 
glimpse(df)

# Save dataset as csv --> 1 row per patient-day (n=15,790 rows, n=1,117 unique patients) 
write_csv(df, here("data", "dm_stg_clean_11.08.21.csv"))

```

# Create condensed dataset where each row represents a unique patient 
```{r Create dataset where each row represents unique pt}
df_unique <- df %>% 
  distinct(ID, .keep_all = T) %>% 
  select(-c(date, death_date, days_to_disc))

# Check to make sure there are n=1,117 rows or unique patients in dataset
glimpse(df_unique)

# Preview final, unique pt dataset 
head(df_unique, n=20)

# Save dataset as csv --> 1 row per patient 
write_csv(df_unique, here("data", "dm_stg_unique_pt_11.08.21.csv"))
```

# End of Document
