---
title: "UCSF Stages Data - Descriptives and Demographics"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project 
MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description
This file summarizes demographic data for COVID positive patients hospitalized at UCSF between [date1-date2] and generates plots to compare length of stay (LOS) by race, ethnicity, and ADI score.

## Source Data
  - Demographics and Daily Covid Stage Data (`dm_covid_adi.csv`)
    + This file contains data on patient demographics and SES-related measures (e.g., ADI, HPI, SVI).
    + Each row contains a unique patient as indicated by their de-identified encounter ID. 

# Load required packages
```{r Load required packages, include=F}
library(here)
library(forcats)
library(tidyverse)
library(labelled)
library(psych)  # describe()
library(tableone)
library(kableExtra)
library(DT)
```

# Read in data 
```{r Import and preview data, echo=F}
df <- read_delim("/Users/sreynolds2/Library/Mobile Documents/com~apple~CloudDocs/Covid Staging/covid-stages/data/dm_covid_adi.csv", delim = "|")
```

# Clean up dataset by selecting necessary vars, renaming variables, and transforming to factor type
```{r}
# Create vector 'vars' to indicate which variables to include in final dataset 
vars <- c("deid_enc_id", "age", "sex", "pat_race", "ethnicity", "BMI", "smoking", "hospital_LOS", "ADI_STATERNK_y")

df <- df %>% 
  select(vars) %>% 
  rename(ID = deid_enc_id,
         race = pat_race,
         LOS = hospital_LOS,
         ADI = ADI_STATERNK_y) %>% 
  mutate(sex = as.factor(sex),
         race = as.factor(race),
         ethnicity = as.factor(ethnicity), 
         smoking = as.factor(smoking),
         ADI = as.factor(ADI))

# Collapse levels for `race`
levels(as.factor(df$race))
df$race <- fct_collapse(df$race, 
                        "Other" = c("Other", "American Indian or Alaska Native"),
                        "Unknown" = c("Unknown", "Declined", "Unknown/Declined"),
                        "Native Hawaiian or Other Pacific Islander" = c("Native Hawaiian or Other Pacific Islander", "Other Pacific Islander"))

# Collapse levels for `ethnicity`
levels(as.factor(df$ethnicity))
df$ethnicity <- fct_collapse(df$ethnicity, 
                             "Unknown" = c("Declined", "Unknown", "Unknown/Declined"))

# Collapse levels for `smoking`
levels(as.factor(df$smoking))
df$smoking <- fct_collapse(df$smoking,
                           "Current Smoker" = c("Current Every Day Smoker", "Current Some Day Smoker", "Light Tobacco Smoker", "Smoker, Current Status Unknown"),
                           "Not Current Smoker" = c("Never Smoker", "Former Smoker", "Passive Smoke Exposure - Never Smoker"),
                           "Unknown" = c("Unknown If Ever Smoked", "NULL"))

# Set variable labels 
df <- df %>% labelled::set_variable_labels(age = "Age (in years)", 
                                           sex = "Biological Sex",
                                           race = "Race", 
                                           ethnicity = "Ethnicity",
                                           smoking = "Smoking Status",
                                           LOS = "Length of Stay (LOS)",
                                           ADI = "Area Deprivation Index (ADI)")

# Check variables labels
var_label(df)

```

# Preview dataset and generate summary statistics 
```{r}
head(df) %>% kbl() %>% kable_material(c("hover", "condensed"))

describe(df) %>% kbl() %>% kable_material(c("hover", "condensed"))

datatable(df)
```

# Create descriptive table of demographics 
```{r}
tab1 <- CreateTableOne(vars = c("age", "sex", "race", "ethnicity", "BMI", "smoking", "LOS", "ADI"), data = df)
print(tab1, showAllLevels = T, nonnormal = "LOS", explain = T, varLabels = T, contDigits = 1, noSpaces = T)

#print(tab1, showAllLevels = T, nonnormal = "LOS", explain = T, varLabels = T, contDigits = 1, noSpaces = T) %>% as.data.frame() %>% kbl() %>% kable_material(c("hover", "condensed"))

#?kableone()
#kableone(tab1)
```

# Create boxplot to compare LOS by race/ethnicity
```{r}
ggplot(df, aes(x = LOS, y = race)) + 
  geom_boxplot()

hist(log(df$LOS))
cor(df$LOS, as.numeric(df$ADI), method = "spearman")

cor(df$LOS, as.numeric(df$ADI), use = "pairwise.complete.obs", method = "spearman")
# use  spearman for non-normally distributed data 

ggplot(df, aes(sample=as.numeric(ADI))) +
  stat_qq() +
  stat_qq_line(color='red') +
  labs(y='LOS')

df$LOS.t <- log(df$LOS)

los_fit <- glm(formula = LOS.t ~ ADI, data = df)
summary(los_fit)


summary(lm(formula = sqrt(LOS) ~ as.numeric(ADI), data = df)
```


