---
title: "Clean and Pre-Process Data -> Patient-Days Table"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

# Background

## Project 
MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description
This file pre-processes the demographics and events and COVID vaccination data and returns a cleaned, expanded dataset with one row per patient-day of hospitalization. Steps include: deselecting unecessary variables, transforming to proper variable type, parsing the admit and discharge date, etc.

## Source Data
  - Demographics and Event (`demographics and event table 4.14.22.csv`)
    + This file contains data on patient demographics (age, sex, race, ethnicity), time of admission and discharge, time of death (if applicable), and hospital length of stay (LOS). 
  - COVID Vaccinations (`COVID vaccinations 4.14.22.csv`)
    + This file contains data on COVID vaccination history, including, dose number, date of vaccination, and manufacturer.
  - All Encounter Diagnoses (`all encounter diagnoses 4.14.22.csv`)
    + This file contains data on comorbid diagnoses. Used to convert to Charlson Comborbidity Index (CCI) score.

# Load required packages 
```{r Load required packages, message = F}
library(here)
library(tidyverse)
library(forcats)
```

# Import data and select necessary variables 
```{r Import and preview data, echo=FALSE}
dm_raw <- read_delim(here("data", "raw", "demographics and event table 4.14.22.csv"), delim = "|")

colnames(dm_raw)

# Identify necessary variables and assign to vector `vars`
vars <- c("deid_enc_id", "deid_pat_id", "age", "sex", "pat_race", "ethnicity", "marital_status", "language", "smoking", "BMI", "covid_pos", "end_in_death", "death_time", "admit_time", "discharge_time", "hospital_LOS", "last_code_status")

dm <- dm_raw %>% 
  select(vars) 
```

# Recode variables to proper format (numeric, factor, or date)
```{r Recode variables to proper format}
# Identify variables to be converted to factor type and assign to vector `vars_fct`
vars_fct <- c("sex", "pat_race", "ethnicity", "marital_status", "language", "smoking", "covid_pos", "end_in_death", "last_code_status")

# Transform variables above to factor type
dm <- dm %>% mutate_at(vars_fct, as.factor)

# Identify variables to be converted to numeric type and assign to vector `vars_num`
vars_num <- c("age", "BMI", "hospital_LOS")

# Transform variables above to numeric type
dm <- dm %>% mutate_at(vars_num, as.numeric)

# Transform admit/discharge/death time variables to date type
dm$discharge_time <- as.POSIXct(dm$discharge_time, format = "%Y-%m-%d %H:%M:%S")
dm$death_time <- as.POSIXct(dm$death_time, format = "%Y-%m-%d %H:%M:%S")
dm$admit_time <- as.POSIXct(dm$admit_time, format = "%Y-%m-%d %H:%M:%S")
```

# Collapse levels for factor variables and set reference level
```{r Collapse levels for factor variables and set reference level, tidy = T, tidy.opts = list(width.cutoff = 60)}
# Collapse levels for race and set ref level 
dm$pat_race <- fct_collapse(dm$pat_race,
                            "Unknown" = c("Unknown", "Declined", "Unknown/Declined", "NULL"),
                            "Other" = c("Native Hawaiian", "Native Hawaiian or Other Pacific Islander", "Other", "Other Pacific Islander", "American Indian or Alaska Native")) %>% 
  relevel(dm$pat_race, ref = "White or Caucasian")

# Collapse levels for ethnicity and set ref level
dm$ethnicity <- fct_collapse(dm$ethnicity, 
                             "Unknown" = c("Unknown", "Declined", "Unknown/Declined", "NULL")) %>% 
  relevel(dm$ethnicity, ref = "Not Hispanic or Latino")

# Collapse levels for smoking and set ref level
dm$smoking <- fct_collapse(dm$smoking,
                           "Current Smoker" = c("Current Every Day Smoker", "Current Some Day Smoker", "Heavy Tobacco Smoker", "Light Tobacco Smoker", "Smoker, Current Status Unknown"),
                           "Not Smoker" = c("Former Smoker", "Never Smoker", "Passive Smoke Exposure - Never Smoker"),
                           "Unknown" = c("Unknown If Ever Smoked", "NULL")) %>% 
  relevel(dm$smoking, ref = "Not Smoker")

# Set ref level for sex 
dm$sex <- relevel(dm$sex, ref = "Male")

# Collapse levels for language and set ref level
dm$language <- factor(ifelse(dm$language=="English", "English", "Not English")) %>% 
  relevel(dm$language, ref = "English")

# Collapse levels for marital_status and set ref level
dm$marital_status <- fct_collapse(dm$marital_status,
                           "Single" = c("Divorced", "Legally Separated", "Single", "Widowed", "RDP-Dissolved", "RDP-LG SEP", "RDP-Widow"),
                           "Paired" = c("Married", "Significant Other", "Registered Domestic Partner"),
                           "Unknown" = c("Unknown/Declined", "NULL")) %>% 
  relevel(dm$smoking, ref = "Single")
```

# Create dataset where each row represent a single patient-day
## Parse date from admit/discharge times and create new variable that spans from admission to discharge for each encounter
```{r Create pt-day table}
dm_clean <- dm %>%
  filter(covid_pos=="Yes" & hospital_LOS>=1) %>%    # Filter for COVID+ pts with LOS>=1
  rename(race = pat_race,
         LOS = hospital_LOS) %>% 
  mutate(admit_date = as.Date(admit_time),          # Parse date from admit/discharge/death times
         discharge_date = as.Date(discharge_time),
         death_date = as.Date(death_time)) %>% 
  group_by(deid_enc_id) %>% 
  # Create new var that spans from admit date to discharge date for each encounter 
  mutate(day_date = list(seq(from = min(admit_date), to = max(discharge_date), by = "day"))) %>% 
  unnest(day_date) %>% 
  relocate(day_date, .after = 2) %>%                # Relocate `day_date` as 3rd col
  select(-c(covid_pos, ends_with("_time")))

# Summarize `dm_clean` 
summary(dm_clean)
```

# Save `dm_clean` as CSV and RDS 
```{r Save dm_clean, eval = F}
write_csv(dm_clean, here("data", "dm_covid_4.14.22.csv"))
saveRDS(dm_clean, here("data", "dm_covid_4.14.22.rds"))
```

# Import COVID vaccinations table and merge with demographics  
```{r Import vaccinations tbl and join with dm tbl}
# Read in covid vaccinations table 
vac <- read_delim(here("data", "raw", "COVID vaccinations 4.14.22.csv"), delim = "|") %>%
  filter(immune_date!="NULL" & dose_num==1) %>% 
  select(-manufacturer)

# Join dm and vac tables on `deid_pat_id`
dm_w_vac <- dm_clean %>% 
  left_join(vac, by = "deid_pat_id") %>% 
  mutate(vac = as.factor(case_when(immune_date < admit_date ~ "Yes", 
                         TRUE ~ "No"))) %>% 
  relocate(vac, .after = "day_date") %>% 
  select(-c(immune_date, dose_num))

# Preview final dataset 
glimpse(dm_w_vac)
```

# Save final `dm_w_vac` dataset as CSV and RDS
```{r Save and export final dataset, eval = F}
write_csv(dm_w_vac, here("data", "dm_covid_4.14.22.csv"))
saveRDS(dm_w_vac, here("data", "dm_covid_4.14.22.rds"))
```

