---
title: "Binary Clinical Scenarios and O2 Device Classifications"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

## Project

MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description

This file imports flowsheet vitals and med admin data and returns a dataset containing all binary clinical scenarios and O2 device usage per patient-day. The resulting dataset will then be converted to ordinal stages of severity, based on WHO Clinical Progression Scale.

## Source Data

-   **Flowsheet Vitals and O2 Device Data** (`flowsheet data table 4.14.22.csv`)
    -   This file contains data on SpO2, FiO2, O2 flow rate, O2 device, ventilation modes and settings, dialysis settings, ECMO settings, etc.
-   **Medical Administration Record** (`med admin table 4.14.22`)
    -   This file contains data on vasopressor (VP) name, VP rate, and VP action.
-   ***Cleaned*** **Demographics and Event Table** (`dm_covid_4.14.22.csv`)
    -   This file contains data on patient demographics, admit date, death date (if applicable), length of stay (LOS), last code status, and COVID-19 vaccination status.
    -   Cleaned dataset obtained from previous script `Clean_and_Patient-Days_Tbl`.

# Load required packages

```{r Load required packages, message = F}
library(here)
library(tidyverse)
```

# Import data

```{r Import all data, echo=F, message=F}
med <- read_delim(here("data", "raw", "med admin table 4.14.22.csv"))
med <- med[-c(nrow(med)-1, nrow(med)), ]

fs <- read_delim(here("data", "raw", "flowsheet data table 4.14.22.csv"))
fs <- fs[-c(nrow(fs)-1, nrow(fs)), ]

dm_covid <- readRDS(here("data", "dm_covid_4.14.22.rds"))
```

# Clean `dm_covid` table (reformat and select necessary variables)

```{r Clean demographics table}
glimpse(dm_covid)

dm <- dm_covid %>%
  select(-c(deid_pat_id, admit_date, discharge_date)) %>% 
  rename(ID = deid_enc_id,
         date = day_date)
```

# Clean `med` table (reformat and select necessary variables)

```{r Clean med adm table, warning = F}
glimpse(med)

med <- med %>%
  rename(ID = deid_enc_id) %>%  
  mutate(date = as.Date(taken_time, format = "%Y-%m-%d"),
         infusion_rate = as.numeric(infusion_rate)) %>% 
  select(ID, taken_time, date, vasopressor_flag, infusion_rate, COVID_tx_flag) 

# Only include encounter IDs that are included in `dm` table 
med <- med[med$ID %in% dm$ID, ]
```

# Clean `fs` table (reformat and select necessary variables)

```{r Clean flowsheet table, warning=F}
glimpse(fs)

# Create not-in operator
`%nin%` <- Negate(`%in%`)

fs <- fs %>% 
  rename(ID = deid_enc_id) %>% 
  filter(!is.na(meas_value) & clinical_concept %nin% c("Respirations", "Urine_output", "Blood_pressure", "Pulse", "Temperature")) %>%
  mutate(date = as.Date(recorded_time, format = "%Y-%m-%d"), 
         meas_val_chr = meas_value,                 # create new var `meas_val_chr` to keep meas_value as character type
         meas_val_num = as.numeric(meas_value)) %>%   
  select(ID, flo_meas_name, meas_val_num, meas_val_chr, recorded_time, date, clinical_concept)

# Only include encounter IDs that are included in `dm` table 
fs <- fs[fs$ID %in% dm$ID, ]

# Check to make sure number of unique encounter IDs in all 3 datasets are equal
  n_distinct(fs$ID)
  n_distinct(dm$ID)
  n_distinct(med$ID)
  # Each dataset has n=813 unique encounter IDs
```

# Create table to indicate whether each patient received vasopressor each day

-   Create new variable `vp_yn` to indicate whether patient received VP at timestamp

    -   If `vasopressor_flag`=Yes & `infusion_rate`\>0, then `vp_yn`=1

-   Group by ID and date, then create variable `VP` to indicate whether pt received VP at any timestamp during day (take max value of `vp_yn`)

-   Remove duplicate rows

```{r Vasopressor}
vp.tab <- med %>% 
  mutate(vp_yn = ifelse((vasopressor_flag == "Yes" & infusion_rate > 0), 1, 0)) %>%
  group_by(ID, date) %>%
  mutate(VP = max(vp_yn, na.rm = T)) %>%
  distinct(ID, date, VP)
```

# Create table to indicate whether each patient received ECMO each day

-   Create new variable `ecmo_yn` to indicate whether patient received ECMO at timestamp

    -   If `clinical_concept`='ECMO_settings', then `ecmo_yn`=1

-   Group by ID and date, then create variable `ECMO` to indicate whether pt received ECMO at any timestamp during day (take max value of `ecmo_yn`)

-   Remove duplicate rows

```{r ECMO}
ecmo.tab <- fs %>% 
  mutate(ecmo_yn = ifelse(clinical_concept == "ECMO_settings", 1, 0)) %>%
  group_by(ID, date) %>%
  mutate(ECMO = max(ecmo_yn, na.rm = T)) %>%
  distinct(ID, date, ECMO)
```

# Create table to indicate whether patient received CRRT each day

-   Create new variable `crrt_yn` to indicate whether patient received CRRT at timestamp

    -   If `clinical_concept`="HD_UF_CRRT_settings" & `flo_meas_name` %in% c("R UFR TRANSCRIBED CRRT IP_CD_UCSF", "R CRRT BLOOD FLOW RATE"), then `crrt_yn`=1

-   Group by ID and date, then create variable `CRRT` to indicate whether pt received CRRT at any timestamp during day (take max value of `crrt_yn`)

-   Remove duplicate rows

```{r CRRT}
crrt.tab <- fs %>% 
  mutate(crrt_yn = ifelse((clinical_concept == "HD_UF_CRRT_settings" &
                              flo_meas_name %in% c("R UFR TRANSCRIBED CRRT IP_CD_UCSF", "R CRRT BLOOD FLOW RATE")), 1, 0)) %>%  
  group_by(ID, date) %>% 
  mutate(CRRT = max(crrt_yn, na.rm = T)) %>%
  distinct(ID, date, CRRT)
```

# Create table to indicate whether patient received support from NIV device each day

-   Create new variable `niv_yn` to indicate whether patient received support from an NIV device at timestamp

    -   If `clinical_concept`="NIV_settings", then `niv_yn`=1

-   Group by ID and date, then create variable `NIV_PER_DAY` to indicate number of times that an NIV device was recorded during day

-   Remove duplicate rows

```{r NIV settings}
niv.tab <- fs %>% 
  mutate(niv_yn = ifelse(clinical_concept == "NIV_settings", 1, 0)) %>% 
  group_by(ID, date) %>% 
  mutate(NIV = max(niv_yn),
         NIV_PER_DAY = sum(clinical_concept == "NIV_settings")) %>% 
  distinct(ID, date, NIV, NIV_PER_DAY)
```

# Create table to indicate whether patient received HD each day

-   Create new variable `hd_yn` to indicate whether patient received support from an HD device or HD settings at timestamp

    -   If `clinical_concept`="HD_UF_CRRT_settings" & `flo_meas_name` %in% c("R HD BLOOD FLOW RATE", "R HD ULTRAFILTRATION RATE"), then `hd_yn`=1

-   Group by ID and date, then create variable `HD` to indicate whether pt received HD support at any timestamp during day (take max value of `hd_yn`)

-   Remove duplicate rows

```{r HD settings}
hd.tab <- fs %>% 
  mutate(hd_yn = ifelse((clinical_concept == "HD_UF_CRRT_settings" & 
                           flo_meas_name %in% c("R HD BLOOD FLOW RATE", "R HD ULTRAFILTRATION RATE")), 1, 0)) %>%
  group_by(ID, date) %>% 
  mutate(HD = max(hd_yn, na.rm = T)) %>% 
  distinct(ID, date, HD)
```

# Create table to indicate whether patient was intubated each day

-   Create new variable `intub_yn` to indicate whether patient was intubated at timestamp

    -   If `clinical_concept`="Intubation_settings" , then `intub_yn`=1

-   Group by ID and date, then create variable `INTUB` to indicate whether pt was intubated at any timestamp during day (take max value of `intub_yn`)

-   Remove duplicate rows

```{r Intubation}
intub.tab <- fs %>% 
  mutate(intub_yn = ifelse(clinical_concept == "Intubation_settings", 1, 0)) %>% 
  group_by(ID, date) %>% 
  mutate(INTUB = max(intub_yn, na.rm = T)) %>% 
  distinct(ID, date, INTUB)
```

# Filter data for SpO2 and FiO2 and merge tables to calculate SF

-   Filter data for SpO2 and FiO2

-   Merge tables and calculate the pulse oximetric saturation SpO2/FiO2 (SF) ratio, a non-invasive surrogate measure for PF ratio

    -   `SF` = (SpO2/FiO2) \* 100

```{r SpO2, FiO2, SF tables}
# Filter data for SpO2 values and assign to table `spo2.tab`
spo2.tab <- fs %>% 
  filter(clinical_concept == "SpO2") %>% 
  select(ID, recorded_time, date, meas_val_num)

# Filter data for FiO2 values and assign to table `fio2.tab`
fio2.tab <- fs %>% 
  filter(flo_meas_name == "R FIO2") %>% 
  select(ID, recorded_time, date, meas_val_num)

# Merge SpO2 and FiO2 tables and calculate SF 
SF.tab <- fio2.tab %>% 
  left_join(spo2.tab, by = c("ID", "recorded_time"), suffix = c(".fi", ".sp")) %>% 
  mutate(SF = (meas_val_num.sp / meas_val_num.fi * 100)) %>% 
  select(ID, recorded_time, date.fi, SF) %>% 
  rename(date = date.fi)
```

# Create table to indicate whether patient experienced SpO2/FiO2\<200 at same timestamp, and assign to new variable `SF_LT_200`

-   Using `SF.tab` created above, filter for patients that have a SF ratio of less than 200

-   Create new variable `sf_lt_200_yn` to indicate whether patient had SF ratio \< 200 at timestamp

-   Group by ID and date, then create variable `SF_LT_200` to indicate whether pt was intubated at any timestamp during day (take max value of `sf_lt_200_yn`)

-   Remove duplicate rows

```{r SpO2/FiO2 < 200, warning = F}
SF200.tab <- SF.tab %>% 
  mutate(sf_lt_200_yn = case_when(SF < 200 ~ 1,
                                  is.na(SF) ~ 0,
                                  TRUE ~ 0)) %>% 
  group_by(ID, date) %>% 
  mutate(SF_LT_200 = max(sf_lt_200_yn, na.rm = T)) %>% 
  distinct(ID, date, SF_LT_200)
```

# Assign all O2 devices to respiratory support categories

Read in `o2_dev_names.csv` which lists all O2 devices and their respective O2 support category. Categories include:

-   NONE: `dev_cat`="no_dev"

-   SIMPLE: `dev_cat`="simple_dev"

-   NIV: `dev_cat`="NIV_dev"

-   IV: `dev_cat`="iv_dev"

-   CPAP: `dev_cat`="cpap_dev"

-   NC (Nasal cannula and high-flow nasal cannula): `nc_dev`=1

```{r Read in `o2_dev_names.csv` and filter by respiratory support category}
# Read in `o2_dev_names.csv` 
o2_dev_names <- read_csv(here("data", "o2_dev_names.csv"))

# NONE
no_dev_names <- filter(o2_dev_names, dev_cat == "no_dev")[[1]]

# SIMPLE
simple_dev_names <- filter(o2_dev_names, dev_cat == "simple_dev")[[1]]

# Non-Invasive Ventilation (NIV) 
niv_dev_names <- filter(o2_dev_names, dev_cat == "NIV_dev")[[1]]

# Invasive (IV)
iv_dev_names <- filter(o2_dev_names, dev_cat == "iv_dev")[[1]]

# CPAP
cpap_dev_names <- filter(o2_dev_names, dev_cat == "cpap_dev")[[1]]

# Nasal cannula or high-flow nasal cannula (HFNC)
nc_dev_names <- filter(o2_dev_names, nc_dev == 1)[[1]]
```

# Create table of O2 devices categorized by respiratory support type to indicate whether patient received respiratory support, and if yes, which type

Each column represents a respiratory support category (as coded in previous step): `NONE`, `SIMPLE`, `NIV`, `IV`, `CPAP`, and `NC`. You can think of these as different types of O2 devices, varying in severity.

```{r O2 device classification table}
device.tab <- fs %>%
  filter(clinical_concept == "O2_device") %>%
  mutate(NOdev_yn = ifelse(meas_val_chr %in% no_dev_names, 1, 0),
         SIMPLEdev_yn = ifelse(meas_val_chr %in% simple_dev_names, 1, 0),
         NIVdev_yn = ifelse(meas_val_chr %in% niv_dev_names, 1, 0),
         IVdev_yn = ifelse(meas_val_chr %in% iv_dev_names, 1, 0),
         CPAPdev_yn = ifelse(meas_val_chr %in% cpap_dev_names, 1, 0),
         NCdev_yn = ifelse(meas_val_chr %in% nc_dev_names, 1, 0)) %>% 
  group_by(ID, date) %>% 
  mutate(NODEV = max(NOdev_yn, na.rm = T),
         SIMPLEDEV = max(SIMPLEdev_yn, na.rm = T),
         SIMPLE_PER_DAY = sum(meas_val_chr %in% simple_dev_names),
         NIVDEV = max(NIVdev_yn, na.rm = T),
         IVDEV = max(IVdev_yn, na.rm = T),
         CPAPDEV = max(CPAPdev_yn, na.rm = T),
         NCDEV = max(NCdev_yn, na.rm = T),
         NC_PER_DAY = sum(meas_val_chr %in% nc_dev_names)) %>% 
  distinct(ID, date, NODEV, SIMPLEDEV, SIMPLE_PER_DAY, NIVDEV, IVDEV, CPAPDEV, NCDEV, NC_PER_DAY)
```

# Create table to indicate whether patients received any O2, low O2, or high O2

-   Categorize O2 usage into any, low, and high:

    -   **O2:** If `meas_val_num`\>0, then `O2`=1

    -   **LowO2:** If `meas_val_num`\>0 & `meas_val_num`\<=12, then `LowO2`=1

    -   **HighO2:** If `meas_val_num`\>12, then `HighO2`=1

```{r Low vs. high O2 flow rate table}
# Create vector for O2 flow rate names
O2_names <- c("R OXYGEN FLOW RATE", "R OXYGEN FLOW RATE 2 IP_CD_UCSF")

# Create table with columns 'O2', 'LowO2', and 'HighO2'
O2.tab <- fs %>% 
  mutate(O2 = ifelse(flo_meas_name %in% O2_names & meas_val_num > 0, 1, 0),
         LowO2 = ifelse(flo_meas_name %in% O2_names & meas_val_num > 0 & meas_val_num <= 12, 1, 0),
         HighO2 = ifelse(flo_meas_name %in% O2_names & meas_val_num > 12, 1, 0)) %>% 
  group_by(ID, date) %>% 
  mutate(O2 = max(O2, na.rm = T),
         LowO2 = max(LowO2, na.rm = T), 
         HighO2 = max(HighO2, na.rm = T)) %>% 
  distinct(ID, date, O2, LowO2, HighO2)
```

# Merge `device.tab` and `O2.tab` to create table to indicate whether patient received support from NC and high O2

-   Merge `device.tab` and `O2.tab` to create `O2_DEV.tab`

-   Create variable `NC_GT_12` to indicate whether patient received support from nasal cannula device AND had high O2 flow rate

    -   If `NCDEV`=1 & `HighO2`=1, then `NC_GT_12`=1

```{r Merge device and O2 tables, create variable NC_GT_12}
# Left_join on 'ID' and 'date' columns
O2_DEV.tab <- O2.tab %>%
  left_join(device.tab, by = c("ID", "date")) %>%          #left join????
  mutate(NC_GT_12 = ifelse(NCDEV == 1 & HighO2 == 1, 1, 0))
```

# Merge all tables to aggregate binary clinical scenarios and O2 device classifications

## List of tables to join include:

-   fs
-   dm
-   vp.tab
-   ecmo.tab
-   crrt.tab
-   niv.tab
-   hd.tab
-   intub.tab
-   SF200.tab
-   O2_DEV.tab

```{r Join all tables}
bin_clin_scen_df <- dm %>% 
  #left_join(fs, by = c("ID", "date"), suffix = c("", ".dupcol")) %>% 
  left_join(vp.tab, by = c("ID", "date"), suffix = c("", ".dupcol")) %>%
  left_join(ecmo.tab, by = c("ID", "date"), suffix = c("", ".dupcol")) %>%
  left_join(crrt.tab, by = c("ID", "date"), suffix = c("", ".dupcol")) %>%
  left_join(niv.tab, by = c("ID", "date"), suffix = c("", ".dupcol")) %>%
  left_join(hd.tab, by = c("ID", "date"), suffix = c("", ".dupcol")) %>%
  left_join(intub.tab, by = c("ID", "date"), suffix = c("", ".dupcol")) %>%
  left_join(SF200.tab, by = c("ID", "date"), suffix = c("", ".dupcol")) %>%
  left_join(O2_DEV.tab, by = c("ID", "date"), suffix = c("", ".dupcol")) %>%
  select(-ends_with(".dupcol")) %>%
  mutate(death = if_else(date == death_date, 1, 0, missing = 0))
```

# IMPORTANT: Some patients (n=30) did not have a separate row on the date of their death

## For these patients, need to bind an additional row where date=death_date, death=1, and stage=10

```{r Create extra row for patients with missing row on date of death, and bind to last row}
bin_clin_scen_df <- bin_clin_scen_df %>%
  filter(end_in_death == "Yes") %>% 
  group_by(ID) %>% 
  filter(death_date != max(date)) %>%    # Identify pts with missing row for day of death
  summarise(date = max(date)) %>%        # Pull out day before death date 
  mutate(date = date + 1,                # Create new row where DATE=DATE+1, DEATH=1, and STAGE=10
         death = 1,
         stage = 10, .) %>% 
   bind_rows(bin_clin_scen_df, .) %>%    # Bind to df 
   arrange(ID)                           # Order by ID 

# Check to make sure num of rows where end_in_death==Yes equals num of rows where DEATH==1
  n_death <- sum(bin_clin_scen_df$death)
  n_end_in_death <- filter(bin_clin_scen_df, end_in_death=="Yes") %>% distinct(ID, .keep_all = T) %>% nrow()
  n_death==n_end_in_death # TRUE 

# From code above, you can filter to get df of pts with missing row for date of death
  # There are n=30 unique encounter IDs that need a row binded to the last date representing their death 
pts_missing_row_for_death <- bin_clin_scen_df %>%
  filter(end_in_death == "Yes") %>% 
  group_by(ID) %>% 
  filter(death_date != max(date))
n_distinct(pts_missing_row_for_death$ID)     # N=30
```

# Save resulting dataframe as CSV and RDS

```{r Export and save data }
write_csv(bin_clin_scen_df, here("data", "binary_clin_scen_df_4.14.22.csv"))
saveRDS(bin_clin_scen_df, here("data", "binary_clin_scen_df_4.14.22.rds"))
```
