---
title: "Generate Ordinal Stages of Severity"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

## Project 
MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description
This file imports `bin_clin_scen_df_4.14.22.rds` which contains all binary clinical scenarios and O2 devices for all patients, and runs a function to generate WHO Clinical Stages of Severity (4-10).

## Source Data
  - Binary Clinical Scenarios and O2 Devices (`bin_clin_scen_df_4.14.22.rds`)
    + This file contains data on patients' demographics, O2 flow rates, clinical scenarios, receipt of O2 devices and their respective category of respiratory support.

# Load required packages 
```{r Load required packages, message = F}
library(here)
library(tidyverse)
```

# Import and preview data 
```{r Import and preview data}
bin_clin_scen_df <- readRDS(here("data", "binary_clin_scen_df_4.14.22.rds"))

# Preview data and variable types
head(bin_clin_scen_df)
glimpse(bin_clin_scen_df)
```

# Create function to get COVID stages based on patients' binary clinical scenarios and usage of O2 devices (and their respiratory support category)
```{r Get stages function}
clin_scen_stg_df <- bin_clin_scen_df %>% 
  mutate(stage = case_when(death==1 ~ 10,
                           ECMO==1  ~ 9,
                           SF_LT_200==1 & (INTUB==1 | IVDEV==1) & (VP==1 | CRRT==1) ~ 9, 
                           VP==1 | CRRT==1 ~ 8,
                           SF_LT_200==1 & (INTUB==1 | IVDEV==1) ~ 8,
                           SF_LT_200==0 & (INTUB==1 | IVDEV==1) ~ 7,
                           (NIVDEV==1 & CPAPDEV==0) & IVDEV==0 & 
                              ((NIV_PER_DAY > NC_PER_DAY) | (NC_PER_DAY>0 & NC_GT_12==1)) ~ 6,
                           NC_PER_DAY>0 & NC_GT_12==1 ~ 6,
                           SIMPLEDEV==1 & (LowO2==1 | HighO2==1 | SIMPLE_PER_DAY>1) ~ 5,
                           NC_PER_DAY>0 & NC_GT_12==0 ~ 5,
                           LowO2==1 & SIMPLEDEV==0 & NIVDEV==0 & IVDEV==0 & CPAPDEV==0 & NCDEV==0 & 
                              INTUB==0 ~ 5,
                           TRUE ~ 4))
```

# Simplify dataset to only include demographics, death, and stage
```{r Filter to create dataset of demographics, death, and stage}
colnames(clin_scen_stg_df)

dm_covid_stg <- clin_scen_stg_df %>% 
  select(-c(end_in_death, death_date:NC_GT_12))

dm_covid_stg
```

# Save and export as CSV and RDS 
```{r Save and export data}
write_csv(clin_scen_stg_df, here("data", "clin_scen_stg_df_4.14.22.csv"))
saveRDS(clin_scen_stg_df, here("data", "clin_scen_stg_df_4.14.22.rds"))

write_csv(dm_covid_stg, here("data", "dm_covid_stg_4.14.22.csv"))
saveRDS(dm_covid_stg, here("data", "dm_covid_stg_4.14.22.rds"))
```

