---
title: "Descriptive Analysis for Covid Staging Data"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project

MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description

This file summarizes demographic and staging data for COVID-19 patients, creates a table of frequency counts, and generates various plots.

## Source Data

-   Demographics and Covid Staging Data (`covid_dm_stages_unique_FINAL_4.14.22.rds`)
    -   This file contains data on patient demographics and COVID stage (based on WHO Clinical Progression Scale, which aims to capture patient clinical trajectory and resource usage over the course of the clinical illness -- in this case, COVID-19).
    -   Each row represents a unique encounter.

# Load required packages and read in data

```{r Load required packages, include=F}
library(tidyverse)
library(psych)
library(tableone)
```

# Read in data and preview

```{r Read in data and preview}
df <- readRDS("/Users/sreynolds2/Documents/GitHub/MS-Covid_Staging/data/covid_dm_stages_unique_FINAL_4.14.22.rds")
df
```

# Summarize data and create Table One

```{r Summarize data and create Table One}
describe(df)

vars <- colnames(df)[2:17]

t1 <- CreateTableOne(data=df, vars=vars)
print(t1, showAllLevels = T)
```

# Figures to summarize data 

```{r}
# Histogram of age

ggplot(df) +
  aes(x = age) +
  geom_histogram(bins = 30L, fill = "#0C4C8A") +
  labs(x = "Age",
       y = "Frequency",
       title = "Distribution of Age") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20L, hjust = 0.5))


# Histogram of BMI

ggplot(df) +
  aes(x = BMI) +
  geom_histogram(bins = 30L, fill = "#0C4C8A") +
  labs(x = "BMI",
       y = "Frequency",
       title = "Distribution of BMI") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20L, hjust = 0.5))


# Histogram of LOS
# Truncated: 0 to 80 days

ggplot(df) +
  aes(x = LOS) +
  geom_histogram(bins = 30L, fill = "#0C4C8A") +
  labs(
    x = "Number of Days",
    y = "Frequency",
    title = "Distribution of Length of Stay (LOS)",
    subtitle = "for patients admitted at moderate",
    caption = "X-axis limits 0 to 80"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  ) +
  xlim(0, 80)


# Histogram of LOS
# Not truncated

ggplot(df) +
  aes(x = LOS) +
  geom_histogram(bins = 30L, fill = "#0C4C8A") +
  labs(x = "Number of Days",
       y = "Frequency",
       title = "Distribution of Length of Stay (LOS)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20L, hjust = 0.5))


# Histogram of LOS for patients admitted at moderate
# Truncated: 0 to 80 days

df %>%
  filter(stage_adm.cat %in% "Moderate") %>%
  ggplot() +
  aes(x = LOS) +
  geom_histogram(bins = 30L, fill = "#0C4C8A") +
  labs(
    x = "Number of Days",
    y = "Frequency",
    title = "Distribution of Length of Stay (LOS)",
    subtitle = "for patients admitted at moderate",
    caption = "X-axis limits 0 to 80"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  ) +
  xlim(0, 80)


# Histogram of LOS for patients admitted at severe
# Truncated: 0 to 80 days

df %>%
  filter(stage_adm.cat %in% "Severe") %>%
  ggplot() +
  aes(x = LOS) +
  geom_histogram(bins = 30L, fill = "#0C4C8A") +
  labs(
    x = "Number of Days",
    y = "Frequency",
    title = "Distribution of Length of Stay (LOS)",
    subtitle = "for patients admitted at severe",
    caption = "X-axis limits 0 to 80"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  ) +
  xlim(0, 80)


# Bar graph of race

df %>%
  ggplot() +
  aes(x = race) +
  geom_bar(fill = "#0C4C8A") +
  labs(y = "Frequency", title = "Distribution of Race") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  )


# Bar graph of ethnicity

df %>%
  ggplot() +
  aes(x = ethnicity) +
  geom_bar(fill = "#0C4C8A") +
  labs(y = "Frequency", title = "Distribution of Ethnicity") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  )


# Bar graph of marital status

df %>%
  ggplot() +
  aes(x = marital_status) +
  geom_bar(fill = "#0C4C8A") +
  labs(x = "Marital Status",
       y = "Frequency",
       title = "Distribution of Marital Status") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  )


# Bar graph of language

df %>%
  ggplot() +
  aes(x = language) +
  geom_bar(fill = "#0C4C8A") +
  labs(x = "Language",
       y = "Frequency",
       title = "Distribution of Language") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  )


# Bar graph of smoking status

df %>%
  ggplot() +
  aes(x = smoking) +
  geom_bar(fill = "#0C4C8A") +
  labs(x = "Smoking Status",
       y = "Frequency",
       title = "Distribution of Smoking Status") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  )


# Bar graph of last code status

df %>%
  ggplot() +
  aes(x = last_code_status) +
  geom_bar(fill = "#0C4C8A") +
  labs(x = "Last Code Status",
       y = "Frequency",
       title = "Distribution of Last Code Status") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 hjust = 0.5)
  )


# Histogram of max stage

ggplot(df) +
  aes(x = max_stage) +
  geom_bar(fill = "#112446") +
  labs(x = "Max Stage",
       y = "Frequency",
       title = "Distribution of Max Stage") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 face = "italic",
                                 hjust = 0.5)
  )


# Histogram of admission stage

ggplot(df) +
  aes(x = stage_adm) +
  geom_bar(fill = "#112446") +
  labs(x = "Admission Stage",
       y = "Frequency",
       title = "Distribution of Severity at Admission") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 face = "italic",
                                 hjust = 0.5)
  )


# Stacked bar chart - proportion of max stage by admission severity

ggplot(df) +
  aes(x = stage_adm, fill = max_stage) +
  geom_bar(position = "fill") +
  scale_fill_hue(direction = 1) +
  labs(x = "Admission Stage",
       y = "Proportion",
       title = "Proportion of Max Stage by Admission Severity") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 face = "italic",
                                 hjust = 0.5)
  )


# Stacked bar chart - frequency of max stage by admission severity

ggplot(df) +
  aes(x = stage_adm, fill = max_stage) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(x = "Admission Stage",
       y = "Frequency",
       title = "Distribution of Max Stage by Admission Severity") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.subtitle = element_text(size = 16L,
                                 face = "italic",
                                 hjust = 0.5)
  )


# Side-by-side boxplot of LOS by admission severity
# Truncated: LOS 0 to 80 days

df %>%
  filter(LOS >= 1L & LOS <= 80L) %>%
  ggplot() +
  aes(x = stage_adm.cat, y = LOS) +
  geom_boxplot(shape = "circle", fill = "#FF8C00") +
  labs(
    x = "Admission Severity",
    y = "Length of Stay (days)",
    title = "Length of Stay by Admission Severity",
    caption = "X-axis limits 0 to 80"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 20L, hjust = 0.5))


# Side-by-side boxplot of LOS by admission severity
# Not truncated

ggplot(df) +
  aes(x = stage_adm.cat, y = LOS) +
  geom_boxplot(shape = "circle", fill = "#FF8C00") +
  labs(x = "Admission Severity",
       y = "Length of Stay (days)",
       title = "Length of Stay by Admission Severity") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20L, hjust = 0.5))


# Stacked bar chart - frequency of max stage by admission severity
# Stages categorized into moderate, severe, and dead

ggplot(df) +
  aes(x = stage_adm.cat, fill = max_stage.cat) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(
    x = "Admission Severity",
    y = "Frequency",
    title = "Distribution of Max Stage by Admission Severity",
    caption = "Stages categorized into moderate (4-5),
    severe (6-9),
    and dead (10).",
    fill = "Max Stage"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.caption = element_text(size = 13L,
                                face = "italic",
                                hjust = 0)
  )


# Stacked bar chart - proportion of max stage by admission severity
# Stages categorized into moderate, severe, and dead

ggplot(df) +
  aes(x = stage_adm.cat, fill = max_stage.cat) +
  geom_bar(position = "fill") +
  scale_fill_hue(direction = 1) +
  labs(
    x = "Admission Severity",
    y = "Proportion",
    title = "Proportion of Max Stage by Admission Severity",
    caption = "Stages categorized into moderate (4-5),
    severe (6-9),
    and dead (10).",
    fill = "Max Stage"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20L,
                              hjust = 0.5),
    plot.caption = element_text(size = 13L,
                                face = "italic",
                                hjust = 0)
  )


# Side-by-side boxplot of death status by age

ggplot(df) +
  aes(x = death, y = age) +
  geom_boxplot(shape = "circle", fill = "#4682B4") +
  labs(x = "Death", y = "Age", title = "Death Status by Age") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20L, hjust = 0.5))


# Stacked bar chart of death status by admission severity category

ggplot(df) +
  aes(x = death, fill = stage_adm.cat) +
  geom_bar() +
  scale_fill_hue(direction = 1) +
  labs(x = "Death",
       y = "Frequency",
       title = "Death Status by Admission Severity",
       fill = "Admission Severity") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20L, hjust = 0.5))
```

