---
title: "UCSF Stages Data - Prelim Analysis"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

## Project 
MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description
This file imports demographic and staging data and returns a dataset indicating each patient's `max_stage`, `date_adm`, `date_disc`, `los`, `stage_adm`, `stage_disc`, `days_to_disc`. This file prepares the staging data for analysis. 

## Source Data
  - Demographics and Daily Covid Stage Data (`dm_covid_stg_11.08.2021.csv`)
    + This file contains data on patient demographics and COVID stage (based on WHO Clinical Progression Scale, which aims to capture patient clinical trajectory and resource usage over the course of the clinical illness -- in this case, COVID-19).
    + Each row represents one day per patient in hospital...

# Load required packages
```{r Load required packages}
library(here)
library(tidyverse)
library(psych)
library(tableone)
library(scales)
library(DescTools) # Winsorized mean -- Winsorize(mean(df$var))
```

# Import and preview data 
```{r Import and preview data, echo=F, warning=F}
#df <- read_csv(here("data", "dm_stg_clean_11.08.21.csv"))

df <- read_csv(here("data", "dm_stg_unique_pt_11.08.21.csv"))

glimpse(df)

head(df, n=20)

psych::describe(df, quant = c(.25, .75))
```

# Create table one for categorical and continuous variables 
```{r warning=F}
# Define categorical and continuous variables
cat_vars <- c("sex", "race", "ethnicity", "smoking", "death", "max_stage", "stage_adm", "stage_disc")
cont_vars <- c("age", "BMI", "LOS")

# tableone::print.CreateCatTable 
t1 <- CreateCatTable(data = df, cat_vars)
print(t1, varLabels = T, showAllLevels = T, digits = 1)
# tableone::summary.CreateCatTable
summary(t1)

# tableone::print.CreateContTable
t2 <- tableone::CreateContTable(data = df, cont_vars)
print(t2, nonnormal = "LOS", digits = 1)
# tableone::summary.CreateContTable
summary(t2)

# Stage transition matrices
# Create categorical table of stage_adm vs. stage_disc
CreateCatTable(strata = 'stage_adm', vars = 'stage_disc', data = df)

# Create categorical table of stage_adm vs. max_stage
CreateCatTable(strata = 'stage_adm', vars = 'max_stage', data = df)

```

# Create histogram of LOS 
```{r}
ggplot(df, aes(x = LOS)) +
  geom_histogram(binwidth = 5) + 
  labs(title = 'Length of Stay',
       subtitle = 'for UCSF Patients',
       x = 'Number of Days',
       y = 'Number of Patients') +
  geom_vline(xintercept = median(df$LOS), color = 'red')
```
# Create barplot showing dstribution of max stages
```{r}
ggplot(df, aes(x = max_stage)) + 
  geom_bar() + 
  labs(title = 'Distribution of Max Stages',
       subtitle = 'at UCSF',
       x = 'WHO Clinical Progression Stage',
       y = 'Frequency') + 
  scale_x_continuous(breaks = 4:10) +
  geom_text(stat = 'count', aes(label = after_stat(count)), size = 3, vjust = -0.5)
```
# Run correlation stats 
```{r}
# max_stage and LOS
cor.test(x = df$max_stage, y = df$LOS, method = 'pearson')

# max_stage and stage_disc
cor.test(x = df$max_stage, y = df$stage_disc, method = 'pearson')

# max_stage and stage_adm
cor.test(x = df$max_stage, y = df$stage_adm, method = 'pearson')

```

# Create proportional stacked bar chart to show proportion of max_stage within stage_adm
```{r}
df %>% 
  group_by(stage_adm, max_stage) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = as.factor(stage_adm), y = count, fill = as.factor(max_stage))) +
  geom_col(position = 'fill') +
  labs(title = 'Distribution of Max Stage by Admission Stage',
       subtitle = 'at UCSF',
       x = 'Admission Stage',
       y = 'Proportion',
       fill = 'Max Stage') #+
  #geom_text(aes(label = stat(y), group = stage_adm), stat = 'summary', fun = sum, vjust = -1)
```

# Create boxplot of LOS grouped by admission stage
```{r}
ggplot(df, aes(x = as.factor(stage_adm), y = LOS)) + 
  geom_boxplot() + 
  labs(title = 'Length of Stay (LOS) by Admission Stage',
       x = 'Admission Stage',
       y = 'Length of Stay')
```
# Transform admission stage, discharge stage, and max stage to stage categories where 
# 4-5 = moderate, 6-9 = severe, and 10 = dead
```{r}
# Collapse levels of `stage_disc` to moderate, severe, and dead --> `stgcat_disc`
df$stgcat_disc <- fct_collapse(as.character(df$stage_disc),
                           "Moderate" = c("4", "5"),
                           "Severe" = c("6", "7", "8", "9"),
                           "Dead" = "10")
df$stgcat_disc <- factor(df$stgcat_disc, levels=c("Moderate", "Severe", "Dead"))
levels(df$stgcat_disc)

# Collapse levels of `stage_adm` to moderate, severe, and dead --> `stgcat_adm`
df$stgcat_adm <- fct_collapse(as.character(df$stage_adm),
                           "Moderate" = c("4", "5"),
                           "Severe" = c("6", "7", "8", "9"),
                           "Dead" = "10")
df$stgcat_adm <- factor(df$stgcat_adm, levels=c("Moderate", "Severe", "Dead"))
levels(df$stgcat_adm)

# Collapse levels of `max_stage` to moderate, severe, and dead --> `stgcat_max`
df$stgcat_max <- fct_collapse(as.character(df$max_stage),
                           "Moderate" = c("4", "5"),
                           "Severe" = c("6", "7", "8", "9"),
                           "Dead" = "10")
df$stgcat_max <- factor(df$stgcat_max, levels=c("Moderate", "Severe", "Dead"))
levels(df$stgcat_max)

```

# Re-run the proportional stacked bar chart and boxplot from above using stage
# categories (moderate, severe, dead)
```{r}
# Proportional stacked bar chart showing which proportion of patients admitted at moderate and severe stage reached a peak of moderate and/or severe stage 
df %>% 
  group_by(stgcat_adm, stgcat_max) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = stgcat_adm, y = count, fill = stgcat_max)) +
  geom_col(position = 'fill') +
  labs(title = 'Distribution of Max Stage by Admission Stage',
       subtitle = 'at UCSF',
       x = 'Admission Stage',
       y = 'Proportion',
       fill = 'Max Stage')

# Boxplot showing median LOS by admission stage category (moderate vs. severe)
ggplot(df, aes(x = stgcat_adm, y = LOS)) + 
  geom_boxplot() + 
  labs(title = 'Length of Stay (LOS) by Admission Stage',
       x = 'Admission Stage',
       y = 'Length of Stay')

```

# End of Document 
